Research and innovation in the smart city is becoming a critical field for our society. Half of the world population is already living in cities, research in smart cities can have a big impact on a considerable fraction of world's population.
Studies demonstrate that social connections in cities stimulates creativity and improves work quality\cite{florida_cities_2005}. This is only one of the reasons why the percentage of people living in urban environments is growing.

Cities are important research playgrounds for substantial improvements in safety and quality of life.
Flow of information and people are easily observable in a highly connected urban environment, at the same time these information affect a high percentage of the humanity. A critical scenario emerges, where decision makers, citizens and researchers can shape the urban environment that influence our society.

It is also important to support the learning process within the urban communities: \textit{lifelong learning} can be used to promote sustainable behaviors and knowledge building in the city, with the ultimate goal of increasing community awareness and participatory creation of the urban space and thus increase quality of citizens' life.

Smart cities present, by definition, a strong technological component.
In Technology Enhanced Learning (TEL), the role of technology is to direct, foster thinking and facilitate the acquisition of higher order skills\cite{goodyear_technologyenhanced_2010}.
Current research applied to learning in the cities seem to focus on two main technological means for learning contents: situated large displays and mobile devices, intended as tablets and smart-phones\cite{luff_mobility_1998}.
% add more ref

Yet the affordances of digital screens are not always suited for informing people in urban settings \cite{koeman}. While large displays suffer from issues known as "display blindess" \cite{}, and in engaging users to interact with them \cite{}; mobile displays (e.g. smartphones, tablets) can be an obstacle to social interaction and are not suitable or safe to be used while in mobility for example walking or biking \cite{}. 

We claim there is a space of opportunities for smart city learning applications in adopting novel ubiquitous computing approaches like tangible interfaces and augmented objects. These technologies have already be found effective in supporting learning\cite{stanton_classroom_2001}, but their applications were mainly oriented to support learning as it happen in conventional schools and classrooms. The principal advantages in adopting these types of interfaces are (i) to enable the creation of rich and unobtrusive user experiences, (ii) to extend the type of data that can be captured to be used as learning content, including sensor data from the environment and from citizens' whereabouts. Therefore sensor-based tangible interfaces could complement traditional approaches based on large screens and smartphones, especially when the learning environment can be as wide and heterogeneous as a city. Todays' increasingly adoption of sensors and IoT technologies are acting as enabling factors for the development of such interfaces; yet whether a number of studies have reported design guidelines for urban screens \cite{}, there a lack of guidelines to help the design of different type of interfaces.

Toolkits play a role for facilitating the design and building prototypes of such applications.

In this paper we...

Our approach is...



to research new forms of utilizing and collecting data in the city and at the same time experimenting with novel strategies of interaction between users and technology.
The potential of custom-designed hardware can open new possibilities in terms of communication efficiency, methods of interaction with the user and sensor data collection.
Glances, gesture based input and micro interactions are some of the approaches possible with the use of latest toolkits for IoT.
%% ref toolkit

Traditional technology is otherwise a limiting factor: mobile devices and large screens support a very strict and confined set of interaction strategies. It's often not possible to tailor the user experience to properly fit the specific scenario because technology is too limiting.
Our goal is to design aiming at the best possible strategy for the users, building the technology around this process and avoiding the constraints typically introduced by more general-purpose hardware/software combinations.

