\subsection{Smart city learning}
As for the concept of smart city, smart city learning is an overloaded concept. Though the term smart city learning has gained popularity, it seems there is no general understanding of the concept. As for many other terms, it does not seem to make sense to aim at a precise definition since its power is in creating an overlying umbrella.

The concept of smart-city has also been used in many different context and is associated with distinctive and innovative aspects that are often quite different. Big diversities are observed on the reasons \textit{why} different cities are defined as \textit{smart}.

This situation is the consequence of the lack of a clear and recognized definition of smart city.

In attempting to pin down what is smart about the smart city, one finds that not only does it involve quite a diverse range of things (information technology, business innovation, governance, communities and sustainability) it can also be suggested that the label itself often makes certain assumptions about the relationship between these things, for example regarding consensus and balance\cite{hollands_will_2008}.

Komninos\cite{komninos_intelligent_2002}, in his attempt to delineate the intelligent city, (perhaps the concept most closely related to the smart city), sees intelligent (smart) cities as ``\textit{territories with high capacity for learning and innovation, which is built\textendash in the creativity of their population, their institutions of knowledge creation, and their digital infrastructure for communication and knowledge management}''.

Smart cities are also a powerful ecosystem for learning. Smart city learning aim to support the improvement of all key factors contributing to the regional competitiveness: mobility, environment, people, quality of life and governance. The approach is aimed at optimizing resource consumption and saving time improving flows of people, goods and data\footnote{http://www.mifav.uniroma2.it/inevent/events/sclo/}.

Education in this context is pursued as a bottom-up process, where person and places are central. Smartness from a learning perspective exists both in the ambient data collected and among the communities that exists within a city.

The separation between student and teacher will fade out. Their role will be content or situation dependent: everybody will be a learner and the relation between persons will get a bigger role.

From the learning perspective, smart cities can be seen as an independent \textit{learning actor} that behaves like an autonomous entity which adapt itself in an evolving environment.

Despite this, the point of view we're more interested in is the one that sees the smart city as a place where citizens learn smart-behaviors.

This type of scenario can possibly involve traditional education which happens in facilities like schools and universities. Our target is instead more narrow and oriented towards \textit{lifelong learning}, defined as the continuous build of skills to adapt and collaborate in dynamic ecosystems like smart cities.


\subsection{Characteristics of Smart City Learning Applications}

%%...what are smart city learning applications?
%%...what kind of learning and what learning goals?

Technologies like mobile devices, tags, web based applications, geographical information and e-learning systems have already been used to develop smart city learning applications on the field\cite{perez-sanagustin_multichannel_2013}\cite{delfatto_geographic_2013}.

%%%%%%
Smart city learning applications consist in the implementation of urban informatics techniques and approaches to promote innovative engagement strategies\cite{amayocaldwell_urban_2013}.
Studies found that urban informatics provide an innovative opportunity to enrich students' place of learning within the city\cite{amayocaldwell_urban_2013}.

No doubt that among the consequences of such attention there is an acceleration in supporting the integration and embedding of ICT within physical environments to realize what has been defined the \textit{everyware}\cite{giffinger_smart_2007}.

This aspect of physicality and situatedness was underlined by other studies, in \cite{giovannella_scenarios_2013} emerges the need for a new technological infrastructure \textit{to foster a more intimate contact with the cultural background of cities/territories and to support living experience characterized by an high level of physical involvement. A kind of involvement very different from staying ``glued'' to a laptop screen or to a portable playstation}.

% Other studies focused on the transformation from a \textit{space} of learning to a \textit{place} of learning in the smart city context, this is achieved through the exposure of real world urban sites and issues\cite{amayocaldwell_urban_2013}.

The use of technology and its ubiquitous nature, has the potential to encourage a wider range of participation through the sharing of information\cite{fredericks_augmenting_2013}.

The learning factor is also complementary to the technology aspect in smart city learning applications.

Tim Campbell, author of Beyond Smart Cities\cite{campbell_smart_2013}, states that learning occurs in the minds of people who are interested in taking action and who care about the cities where they live.

We identified a few learning approaches that can be considered relevant for smart city learning.

Mobile learning is \textit{any sort of learning that happens when the learner is not at a fixed, predetermined location, or learning that happens when the learner takes advantage of the learning opportunities offered by mobile technologies} \cite{omalley_guidelines_2005}. Ubiquitous learning on the other hand means something more than mobility, a ubiquitous learning environment is any setting in which students can become totally immersed in the learning process \cite{syvanen_supporting_2005}; it is a setting of pervasive learning that is happening all around the student but the student may not even be conscious of the learning process. In extension, smart city learning leverages the infrastructure and services that modern smart cities offer in order to transform them into open learning environments\cite{christopoulou_learning_2013}.

As younger generations are ``born'' familiar with those systems there is an increasing benefit in adopting such technologies, as they facilitate students to continue their learning process outside classrooms, when and where they desire, through exploration and interaction. Therefore, an increasing interest in physical spaces, in the role of technology in city-wide environments and in passage of learning activities outside the classroom is revealed\cite{christopoulou_learning_2013}.

From the literature then, some important factors and activities emerges as critical in smart city learning scenarios:
\begin{itemize}
\item urban informatic as opportunity for learning in the city
\item embedding ICT in the physical environment, promoting also physical involvement
\item ubiquitous technology has the potential to encourage participation
\item promoting participation, interest in taking action and involvement also promotes learning
\item ubiquitous, open learning in smart cities can be unconscious and happens through transformation of infrastructures and services
\item interest is increasing in learning connected to physical spaces outside the classrooms and technology
\end{itemize}

%%%%%%
Some of these critical factors in smart city learning scenarios can be addressed with the help of TUIs. We identified some overlapping between the strengths of TUIs and the challenges of SCL scenarios.

The following peculiarities of TUIs can support and empower SCL applications.

\subsubsection{Data collection}
Sensor data can be captured embedding proper electronics in augmented objects and public or private spaces.
Some of this \textit{sensing} capabilities (e.g. accelerometers, gyro) are necessary to actually allow sensor-based user interactions; for example detecting the presence of public in a space and react with an appropriate feedback. Besides this, raw data sensed can be manipulated and processed also for other purposes more connected to the application logic rather than remaining limited to only sense user interaction.
The opportunity to flexibly and quickly prototype custom devices offered by todays' tools such as Arduino [REF], allow to quickly explore multiple design choices adopting several different sensors.
The opportunity to situate sensing capabilities both in the environment and physically on the users widens the domain where data collection take place.

Mobility of sensors embedded in TUIs vs city infrastructure

\subsubsection{Data visualization}
Citizens everyday living the city can be subjected to common life situations where the most suitable interfaces for accessing information can be fairly different. For example train schedule information can be provided by a screen-based interface when at home or by a wearable interface while riding a bike to the station. The capability of delivering the right content at the right time is important because it could improve learning efficacy [REF?]. Tangible interfaces can deliver information using non distracting interfaces such as haptic or ambient feedbacks that might be more appropriate than screens when on the go. that can shuttle simple information to the user in a very fast and effective way.
Simple output components can be combined to implement custom-designed and tailored output primitives. For example a custom shaped led matrix.

\subsubsection{Data processing} %How data processing is used for SCL applications?
Processing of sensed data can introduce a variety of challenges and demands for compromises. Being TUIs connected devices by nature, two main alternatives are feasible, depending on the nature of data and the processing power available. Data can be either processed on-board, transmitted to an external service for storing and/or processing. An hybrid approach can also be used, where data is pre-elaborated or filtered on board and then transmitted for further processing.
We can then state that the best strategy is influenced by:
\begin{itemize}
    \item characteristics of data stream
    \item complexity of the processing routine
    \item processing power available
    \item network availability and capabilities
\end{itemize}

\subsubsection{User interaction} %this overlaps with sec 2.1.2
In many smart city scenarios, users cannot always devote their full attention to interact with devices. TUIs helps at this regard allowing several interaction strategies that can be less distracting and immersive when the scenario does not require it.
Background and foreground interaction methods are both implementable, as well as multi-modal strategies where users can interact via different channels at the same time. Interaction can also be distributed in the space, involving actions performed on/with different augmented objects situated in the space.

\subsubsection{Sharing of artifacts and information}
Technologies like big screens and mobile devices, often used in smart city learning scenarios, present some constraints that are quite rigid and difficult to overcome. For example mobile devices are proven to support cooperative activities and information sharing, but they physically remain strictly personal apparatus that are not comfortably shared even among friends and relatives.
Public displays are instead useful only to present information that is relevant for a defined geographical place, usually where the display itself is located. The nature of information should also be generic enough to not introduce privacy concerns.
With TUIs it is possible to fill the gap between public and personal/private. Devices can be tailored and adapted to support use cases where the users are comfortably sharing objects and informations.
The added flexibility allow for several level of experience that can freely space between the \textit{private} and the \textit{public}.

\subsubsection{Learning, smart cities and ubiquitous}
Tools, as well as ICT applications and the Internet itself, may facilitate higher order skills and thus mediate learning, in schools and beyond\cite{kashtan_outdoors_2013}. Indeed, Engestrom\cite{engestrom_non_1991} argues against the separation between schooling and other learning experiences; he refers to it as the encapsulation of school learning. To overcome this encapsulation, he offers emphasis on the role of mediating artifacts in human cognition and learning.
This project achieves overcoming the encapsulation of school learning by using ICT tools to mediate the environment to the students, as well as to expose the products to the community\cite{kashtan_outdoors_2013}.

In \cite{luff_mobility_1998} an analysis of technology-supported coordination mechanisms is provided for several scenarios.
After investigating a urban scenario in London's underground they reported that:

\textit{For example, in distributing technologies around the environment developing support for station supervisors would appear to be a prototypical case of "ubiquitous computing" or
"augmented reality". However, although typical developments in these areas aim to support tasks and activities by augmenting everyday artifacts with computational capabilities it is not all that clear which artifacts are most relevant for such enhancement or what capabilities should be augmented.}

It is also mentioned that this type of technologies are able to support coordination mechanisms demonstrated to be essential for the scenarios examined, like micro-mobility, ubiquitous information retrieval and glances:

\textit{
In considering the design of the technology however, we have increasingly realised that it may well be a mistake to place access to all necessary resources on the device itself. It is critical for example that the system remains portable, and becomes one of the various tools that station staff carry about with them as part of their normal duties.}

\textit{Nevertheless, it appears from the studies considered in this paper that the micro-mobility of objects may be critical when considering how to support co-present, collaborative activities. To provide for this may require not only both mobile and fixed devices, but quite novel support for mobility that focuses on the moment-to-moment manipulation of objects.}

